---
title: "Arabidopsis - regressions"
author: "Nelson Nazzicari"
date: "`r Sys.Date()`"
output: 
  html_document:
    css: CREA.css
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, echo=FALSE, include=FALSE, message=FALSE}
library(knitr)
options(knitr.kable.NA = ' ', digits = 4) #this is for kable
library(ggplot2)
library(plyr)
library(printr)
library(DT)
library(xfun)
library(tidyr)
library(madbito)
xfun::pkg_load2(c('base64enc', 'htmltools', 'mime'))
```

```{r cheatsheet, echo=FALSE, include=FALSE, eval=FALSE}
# EVALUATE CODE?  eval=TRUE
# PRINT CODE?     echo=TRUE
# PRINT OUTPUT?   include=TRUE
# UNWANTED STUFF? message=FALSE warning=FALSE error=FALSE

#attach a file (e.g. an xls)
xfun::embed_file('some_file.xls')

#data table with fancy selectors and searches
DT::datatable(cars)

#more humble data table
kable(cars, caption = "my cars")
```

```{r echo=FALSE, include=TRUE}
#this is your standard R chunk
```

# Document history

<!--
To highlight changes use
<span class="new"> your stuff </span>
-->

* 2024-03-25 - initial report
* 2024-03-26 - added kinship section
* 2024-04-30 - added section on experiment 2
* 2024-05-06 - added section on SNP based rrBLUP

# Aim
Measure standard genomic regression performances on an arbidopsis dataset.
The code for regressions and report available at: https://github.com/ne1s0n/crea-za_2024_geno_pheno_arabidopsis

# Dataset

* SNP markers: (circa) 1000 markers, subset from 500k 
* traits: Cauline Leaf number (CL), DaysToFlowering (1, 2, 3), Rosette Leaf number (RL)
* accessions: (circa) 1000 accessions, already split in train (80%) and test (20%)

# Regression - experiment 1

* **mode 1 - CROSSVALIDATION** The train set is further split in a 10-folds crossvalidation scheme, and
  always tested on the test set (representing 20% of the full dataset)
* **mode 2 - FULL** The full train set is used for train (once). The remaining test set is used for validation 

## Results

```{r echo=FALSE, include=TRUE}
#loading result data
res = read.csv('/home/nelson/research/crea-za_2024_geno_pheno_arabidopsis/results/accuracy.csv', stringsAsFactors = FALSE)

#averaging over repetitions
res = ddply(res, .(dataset.train, dataset.test, regressor), function(x){
  return(data.frame(
    coeff_det = mean(x$coeff_det),
    pearson = mean(x$pearson),
    rmse = mean(x$rmse)
  ))
})

#removing useless results
res = subset(res, !grepl(pattern = 'train [CV]', x=res$dataset.test, fixed = TRUE))

#screen labeling: mode
res$mode = 'CV'
sel = grepl(pattern = 'FULL', res$dataset.test)
res[sel, 'mode'] = 'FULL'
res$dataset.test = NULL

#screen labeling: trait
res$dataset.train = gsub(pattern = 'CVtrain', replacement = '', res$dataset.train)
res$dataset.train = gsub(pattern = 'FULLtrain', replacement = '', res$dataset.train)
res = col.rename(res, oldname = 'dataset.train', newname = 'trait')

#wide table
res = pivot_wider(res, id_cols = c('trait', 'regressor'), names_from = 'mode', values_from = c('pearson', 'rmse', 'coeff_det'))

#separating kinship and SNP regressors
kinship_selector = grepl(pattern = 'KIN', res$regressor)
res$regressor = gsub(pattern = 'KIN ', replacement = '', res$regressor)

#list of metrics, easier to put it in an array for further formatting
metrics = setdiff(colnames(res), c('trait', 'regressor'))
```


### Cauline Leaf number (CL)
```{r echo=FALSE, include=TRUE}
#showing the table
#c('Cauline Leaf number (CL)', 'DaysToFlowering1', 'DaysToFlowering2', 'DaysToFlowering3', 'Rosette Leaf number (RL)')
sel = res$trait == 'Cauline Leaf number (CL)'
DT::datatable(res[sel,], rownames = FALSE, options = list(dom = 't')) %>% formatRound(metrics, digits=4)
```

### DaysToFlowering1
```{r echo=FALSE, include=TRUE}
#showing the table
#c('Cauline Leaf number (CL)', 'DaysToFlowering1', 'DaysToFlowering2', 'DaysToFlowering3', 'Rosette Leaf number (RL)')
sel = res$trait == 'DaysToFlowering1'
DT::datatable(res[sel,], rownames = FALSE, options = list(dom = 't')) %>% formatRound(metrics, digits=4)
```

### DaysToFlowering2
```{r echo=FALSE, include=TRUE}
#showing the table
#c('Cauline Leaf number (CL)', 'DaysToFlowering1', 'DaysToFlowering2', 'DaysToFlowering3', 'Rosette Leaf number (RL)')
sel = res$trait == 'DaysToFlowering2'
DT::datatable(res[sel,], rownames = FALSE, options = list(dom = 't')) %>% formatRound(metrics, digits=4)
```

### DaysToFlowering3
```{r echo=FALSE, include=TRUE}
#showing the table
#c('Cauline Leaf number (CL)', 'DaysToFlowering1', 'DaysToFlowering2', 'DaysToFlowering3', 'Rosette Leaf number (RL)')
sel = res$trait == 'DaysToFlowering3'
DT::datatable(res[sel,], rownames = FALSE, options = list(dom = 't')) %>% formatRound(metrics, digits=4)
```

### Rosette Leaf number (RL)
```{r echo=FALSE, include=TRUE}
#showing the table
#c('Cauline Leaf number (CL)', 'DaysToFlowering1', 'DaysToFlowering2', 'DaysToFlowering3', 'Rosette Leaf number (RL)')
sel = res$trait == 'Rosette Leaf number (RL)'
DT::datatable(res[sel,], rownames = FALSE, options = list(dom = 't')) %>% formatRound(metrics, digits=4)
```

# Regression experiment 2: hand-picked CV

This experiment was done with the following:

* 10-folds CV covering the whole dataset (coordinated with the UNIBA group, they provided the list of samples for each fold)
* only kinship-based regressors: GBLUP and RKHS (Bayesian GBLUP)
* kinship computed following the [Astle&Balding method](https://www.jstor.org/stable/25681325) using the **full** dataset of half million SNPs
* the whole process repeated five times for numerical stability, presented results are averaged

To this we added another regression, on SNP data, using standard ridge regression (rrBLUP):

* same hand-picked CV as above
* model trained only on rectangular SNP matrix from training set, then applied to test set
* label "SNP rrBLUP"

```{r echo=FALSE, include=TRUE}
#loading result data
res = rbind(
  read.csv('/home/nelson/research/crea-za_2024_geno_pheno_arabidopsis/results/accuracy_regression2_coordinated_CV.csv', stringsAsFactors = FALSE),
#  read.csv('/home/nelson/research/crea-za_2024_geno_pheno_arabidopsis/results/accuracy_regression2_coordinated_CV_kinshipTrainOnly.csv', stringsAsFactors = FALSE),
  read.csv('/home/nelson/research/crea-za_2024_geno_pheno_arabidopsis/results/accuracy_regression2_coordinated_CV_SNP.csv', stringsAsFactors = FALSE)
)
  
#removing info on split
res$dataset.train = gsub(pattern = ' split.*', replacement = '', x = res$dataset.train)

#display names
res$regressor = gsub(pattern = '^rrBLUP$', replacement = 'SNP rrBLUP', res$regressor)

#averaging over repetitions
res = ddply(res, .(dataset.train, regressor), function(x){
  return(data.frame(
    coeff_det = mean(x$coeff_det),
    pearson = mean(x$pearson),
    rmse = mean(x$rmse)
  ))
})


#screen labeling: trait
res = col.rename(res, oldname = 'dataset.train', newname = 'trait')

kable(res)

#wide table
#res = pivot_wider(res, id_cols = c('trait', 'regressor'), names_from = 'mode', values_from = c('pearson', #'rmse', 'coeff_det'))

#separating kinship and SNP regressors
#kinship_selector = grepl(pattern = 'KIN', res$regressor)
#res$regressor = gsub(pattern = 'KIN ', replacement = '', res$regressor)

#list of metrics, easier to put it in an array for further formatting
#metrics = setdiff(colnames(res), c('trait', 'regressor'))
```
